InnoDB Full text Search hacking with Shiny/R

### 2015-12-27 article.rmd
fixes

As we can observ on the previous examples, the query extension algorithm has some noisy produced by the double query
generated in order to join  the results. Also, it won't matter if you have the exact match twice, as the algorithm does not
focus on the exact matches, instead it will relate the occurrences on the first search (where the match happens) and the
sencond (where the algorithm searches for the words that appear in the same phrase than the exact match). That means 
that it is possible to have a high rank without have any exact matching occurrence.

@@

Even worst when using large sentences. In the example bellow you will see the same query, against the table 
storing by paragraph. The boolean rank shows some of the entries way above others, however the query extension
locates at the top records that not necessarily has a lot of exact matches.

@@

node1 [localhost] {msandbox} (test) > set global innodb_ft_aux_table = 'test/bookContent';
Query OK, 0 rows affected (0,00 sec)

node1 [localhost] {msandbox} (test) > SELECT FTS_DOC_ID, group_concat(it.POSITION) as positions, round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank, round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, length(content) as len FROM bookContent bl join information_schema.INNODB_FT_INDEX_TABLE it ON (bl.FTS_DOC_ID = it.DOC_ID)  WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY QERank DESC LIMIT 10 ;
+------------+------------------+--------+-------------+------+
| FTS_DOC_ID | positions        | QERank | BooleanRank | len  |
+------------+------------------+--------+-------------+------+
|        757 | 6858             |   2804 |           1 | 8101 |
|        735 | 2842             |    925 |           1 | 3133 |
|        157 | 747              |    698 |           1 | 2385 |
|        733 | 265,1342,816,750 |    569 |           2 | 1765 |
|        739 | 1171             |    540 |           1 | 1659 |
|        249 | 197,739          |    435 |           1 | 1368 |
|        279 | 664              |    428 |           1 | 1365 |
|        411 | 468,410          |    396 |           2 | 1170 |
|        155 | 31,495,158,525   |    363 |           5 | 1378 |
|        327 | 60               |    362 |           1 | 1372 |
+------------+------------------+--------+-------------+------+
10 rows in set (4,09 sec)
```

@@

sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength 
            FROM bookContent bp  
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)

sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength 
            FROM  bookContentByLine bl 
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)
```

@@

The aggregations shown on the results are over the results using the selected algorithm. The aggregations over
the query expansion ranks are only against the results, not all the rows.

You can start trying `-country +president` in boolean mode.

```{r, echo=FALSE}
        ui <- shinyUI(fluidPage(
                mainPanel(  
                          selectInput("ftsMode", label = "Algorithm:",
                                      choices = c("WITH QUERY EXPANSION", "IN BOOLEAN MODE", "IN NATURAL LANGUAGE MODE"), selected = "WITH QUERY EXTENSION"),
                          textInput("filtering", "FTS Filter:", "default"),
                          submitButton("Stats!"),
                          # uiOutput('tableStat'),
                          uiOutput('table')
                          )
                       ))

@@

>>>>

The noise generated by the query expansion is expected and described in the official documentation [here.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-query-expansion.html)

@@

This is even worser when using large sentences. In the example bellow you will see the same query, against the table 
storing by paragraph. The boolean rank shows some of the entries way above others, however the query extension
locates at the top records that not necessarily has a lot of exact matches.

@@

(test) > set global innodb_ft_aux_table = 'test/bookContent';
Query OK, 0 rows affected (0,00 sec)

(test) > SELECT FTS_DOC_ID, 
                group_concat(it.POSITION) as positions, 
                round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank, 
                round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, 
                length(content) as len 
          FROM bookContent bl join information_schema.INNODB_FT_INDEX_TABLE it ON (bl.FTS_DOC_ID = it.DOC_ID)  
          WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) 
              AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY QERank DESC LIMIT 10 ;
+------------+------------------+--------+-------------+------+
| FTS_DOC_ID | positions        | QERank | BooleanRank | len  |
+------------+------------------+--------+-------------+------+
|        757 | 6858             |   2804 |           1 | 8101 | <<< This one on the top does only 1 exact match.
|        735 | 2842             |    925 |           1 | 3133 |
|        157 | 747              |    698 |           1 | 2385 |
|        733 | 265,1342,816,750 |    569 |           2 | 1765 |
|        739 | 1171             |    540 |           1 | 1659 |
|        249 | 197,739          |    435 |           1 | 1368 |
|        279 | 664              |    428 |           1 | 1365 |
|        411 | 468,410          |    396 |           2 | 1170 |
|        155 | 31,495,158,525   |    363 |           5 | 1378 |  <<< This doc contains 5 exact matches!
|        327 | 60               |    362 |           1 | 1372 |
+------------+------------------+--------+-------------+------+
10 rows in set (4,09 sec)
```

@@

sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength, count(*) as numEntries 
            FROM bookContent bp  
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)

bookContentStats <- rs

sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength, count(*) as numEntries 
            FROM  bookContentByLine bl 
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)

bookContentByLineStats <- rs

@@

The aggregations shown on the results are over the results using the selected algorithm.

You can start trying `+country +(>house <president) -home` in boolean mode. For more information about operator in Innodb's FTS,
you can read [this.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-boolean.html) Using boolean formulas in NATURAL LANGUAGE MODE,
does not have the desired effect, however you'll be able to see how it affects the ranking compared with the boolean search.

NOTE: distance operator not yet supported.

```{r, echo=FALSE}
        ui <- shinyUI(fluidPage(
                mainPanel(  
                          selectInput("ftsMode", label = "Algorithm:",
                                      choices = c("WITH QUERY EXPANSION", "IN BOOLEAN MODE", "IN NATURAL LANGUAGE MODE"), selected = "WITH QUERY EXPANSION"),
                          textInput("filtering", "FTS Filter:", "default"),
                          submitButton("Stats!"),
                          # uiOutput('tableStat'),
                          uiOutput('table')
                          )
                       ))

@@

## Recommendations

- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
  size is crucial when using this.
- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
  single fields. 
- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the 
  FTS functions to avoid double executions. You can do this using somehting like:
  `match(content,title) against ("first (<second >third)") > 1 `
  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches. 
- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
  maintance on the table.
- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.

---------------------- FROM HERE ONLY TRASH TO BE REMOVED -----------------------------
